Isomorphic Interpreters from Logically Reversible Abstract Machines

### 2012-02-24 iso-int/rc-iso-int.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What are the high level steps in the translation?}
\label{sec:steps}

TODO

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What abstract machines can be turned into isomorphic interpreters?}

Logically reversible abstract machines with distinguishable start and
stop states can be converted into isomorphic interpreters in
{{langRevT}}.  To develop an understanding for what sorts of machines
these are, let us look at some examples.

@@

Consider a machine that has the addition rule {{<''x+y''>|--><''z''>}}
where {{z}} is the sum of {{x}} and {{y}}. This machine again cannot
be turned into an isomorphic interpreter because the one reduction
rule does not have a logically inverse -- in other words, given a
{{z}}, we are unable to determine which {{x}} and {{y}} it originated
from.

@@

\noindent
Conditions that must be satisfied for constructing a isomorphic
interpreter:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adder and Multiplier}

\subsection{Logically Reversible Adder}

%subcode{bnf} include main
% Expressions, e = n
% Numbers, n, m,p = 0 | n + 1
%
% Machine states = <e, e, e>
% Start state = <n, p, 0>
% Stop State = <0, p', m>

Small-step Operational semantics:

>>>>

The construction of the isomorphic interpreter essentially exploits
the structure of most abstract machines -- the LHS of the transition
rules determines a machine state by pattern machine. The RHS does the
dual namely term reconstruction. The fact that the machine is
logically reversible ensures that the computation performed by each
reduction is essentially captured by an isomorphism. 
%% To develop an
%% understanding for what sort of machines can be treated this way, let
%% us look at some examples.

@@

Consider a machine that has the addition rule {{<''x+y''>|--><''z''>}}
where {{z}} is the sum of {{x}} and {{y}}. This machine again cannot
be turned into an isomorphic interpreter because the one reduction
rule does not have a logically inverse -- in other words, given a
{{z}}, we are unable to determine the original {{x}} and {{y}}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{What abstract machines can be turned into isomorphic interpreters?}

\paragraph*{Conditions that must be satisfied for constructing a isomorphic
interpreter}:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{High level steps in the translation:}
\label{sec:steps}

\item Flip the construction inside out introducing a {{trace}} and
  exposing the start and stop states.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Adder and Multiplier}

\section{Logically Reversible Adder}

Consider

\begin{multicols}{2}  
%subcode{bnf} include main
% Numbers, n, m,p = 0 | n + 1
% Machine states = <e, e, e>

%subcode{bnf} include main
% Start state = <n, p, 0>
% Stop State = <0, p', m>
\end{multicols}

%Small-step Operational semantics:
